---
title: "Assignment1"
output: html_document
---
# Reproducible Research Peer Assignment-1

## Part 1(mean total number of steps taken per day):
```{r,echo=TRUE}
library(graphics)
library(plyr)
##importing data:
data<-read.csv("./data/activity.csv")
summary(data)

NumberofNA=sum(is.na(data))

##taking sum:
data2<-ddply(data,.(date),summarise, Total=sum(steps))

hist(data2$Total,breaks=7,col="GREEN",main="Histogram of steps taken every day",xlab = "Number of Steps")

##MEDIAN OF THE DATA EXCLUDING THE NA:
median(data2$Total,na.rm=TRUE)

##MEAN OF THE DATA EXCLUDING THE NA:
mean(data2$Total,na.rm=TRUE)

```


# Part 2(Average Daily Activity Pattern):
```{r,echo=TRUE}
data3<- ddply(data,.(interval),summarise,Mean = mean(steps,na.rm=TRUE))
plot(data3$interval,data3$Mean,type='l',xlab="Interval",ylab="Mean")

max<-max(data3$Mean)
maxInterval<-data3[data3$Mean==max,]$interval
```
The 5 minute interval which contains the maximum number of steps taken: `r maxInterval` and the max is:`r max`. The max interval can be interpreted as the time between 8:30-8:35.


# Part 3(Imputing missing values)
```{r,echo=TRUE}
##Replacing the NA values with the relavent Mean.
data[is.na(data$steps),]$steps = data3$Mean
##There are no NA values after replacing NA's.
summary(data)

data2<-ddply(data,.(date),summarise, Total=sum(steps))
hist(data2$Total,breaks=7,col="GREEN",main="Histogram of steps taken every day",xlab = "Number of Steps")
median(data2$Total)
mean(data2$Total)
```
The difference in mean and median due to imputing is pretty less and the number of NA values were `r NumberofNA` before imputing.

#Part 4(differences in activity patterns between weekdays and weekends)

```{r,echo=TRUE}
day<-weekdays(as.Date(data$date))
data<-cbind(data,day)
head(data)
##Construction of the factor:
x <-data$day
x<-as.character(x)
x[x == "Friday"|x== "Tuesday"|x== "Wednesday"|x== "Thursday"|x== "Monday"] <- "Weekday"
x[x=="Saturday"|x== "Sunday"] = "Weekend"
x<-as.factor(x)
weekFactor<-x
is.factor(weekFactor)
data<-cbind(data,weekFactor)
#indicating whether the day is a weekday or a weekend:
head(data)
tail(data)


data5<-ddply(data,.(interval,weekFactor),summarise,Mean = mean(steps))

library(ggplot2)
qplot(interval,Mean,data=data5,facets=weekFactor~.,geom="line")

```


